---
title: "NFSC-CPR"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R package facilitates access and analysis of [NFSC](https://www.fisheries.noaa.gov/about/northeast-fisheries-science-center) [Continuous Plankton Recorder](https://en.wikipedia.org/wiki/Continuous_Plankton_Recorder) (CPR) data.  Data can imported from distrubuted [Excel](https://en.wikipedia.org/wiki/Microsoft_Excel) spreadsheets into data frames, [tibbles](https://tibble.tidyverse.org/) or [sf tables](https://r-spatial.github.io/sf/).

## Requirements

+ [R v4.1+](https://www.r-project.org/)
+ [rlang](https://CRAN.R-project.org/package=rlang)
+ [dplyr](https://CRAN.R-project.org/package=dplyr)
+ [readxl](https://CRAN.R-project.org/package=readxl)
+ [readr](https://CRAN.R-project.org/package=readr)

## Installation

```
remotes::install_github("BigelowLab/nfsccrp")
```

## Usage

### Data path

The data is quiet sizeable to store in a package, so we provide a means to set the data location outside of the package in a way that persists between R sessions.  You can set that path once and then forget it (or you can change it at anytime.)

```{r}
library(nfsccpr)

nfsccpr::set_path("/Users/ben/Library/CloudStorage/Dropbox/data/noaa/nmfs/cpr")
```


Presumably, you have placed a data file in that directory.

### List files in data path

```{r}
ff = nfsccpr::list_files(full.names = TRUE)
ff
```

### Read a file

Reading the file yields once element for each included sheet.

```{r}
x = import_cpr(ff[1])
x
```