---
title: "NFSC-CPR"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R package facilitates access and analysis of [NFSC](https://www.fisheries.noaa.gov/about/northeast-fisheries-science-center) [Continuous Plankton Recorder](https://en.wikipedia.org/wiki/Continuous_Plankton_Recorder) (CPR) data.  Data can imported from distrubuted [Excel](https://en.wikipedia.org/wiki/Microsoft_Excel) spreadsheets into data frames, [tibbles](https://tibble.tidyverse.org/) or [sf tables](https://r-spatial.github.io/sf/).

>Part of this dataset was obtained from [NOAA's NFSC](https://www.fisheries.noaa.gov/about/northeast-fisheries-science-center) on 20 January 2024 and might not represent the most up to date data. Please contact NOAA directly for the most recent data. The balance of this dataset is avaiable from the [NERACOOS ERDDAP server](https://www.neracoos.org/erddap/index.html).

## Requirements

+ [R v4.1+](https://www.r-project.org/)
+ [rlang](https://CRAN.R-project.org/package=rlang)
+ [dplyr](https://CRAN.R-project.org/package=dplyr)
+ [readxl](https://CRAN.R-project.org/package=readxl)
+ [stringr](https://CRAN.R-project.org/package=stringr)
+ [readr](https://CRAN.R-project.org/package=readr)

## Installation

```
remotes::install_github("BigelowLab/nfsccrp")
```

## Usage

### Read the data

The dataset contains both zooplankton and phytoplankton data from the Mid-Atlantic Bight and Gulf of Maine regions.  We regularily access to the zooplankton data, but it is easy to switch to phytoplankton.  We can read the data in its "raw" state as a simple table or transformed into a [sf]() POINT table.

####  Read as a wide table
```{r}
library(nfsccpr)
x = read_cpr(name = "zooplankton", form = "table") |>
  dplyr::glimpse()
```


#### Read as a sf POINT table
```{r}
x = read_cpr(name = "zooplankton",  form = "sf") |>
  dplyr::glimpse() 
```


### Import a raw data file
The NFSC data were shared as a large two-sheet Excel document. It's difficult to know if the format of the raw data file will remain as that distributed in January 2024.  Importing the one we were provided is encoded in a package function.  The package can be adapted for other formats but may require a small effort in adapting the code.

```{r}
ff = "/Users/ben/Library/CloudStorage/Dropbox/data/noaa/nmfs/cpr/MAB CPR (Jan 20, 2024 update).xlsx"
x = import_cpr(ff[1])
x
```

THe GMRI data files are formatted in long CSV files.